# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
dt <- read.csv("activity.csv")
library(reshape2)
dt_melt <- melt(dt, id=c("date", "interval"), meas.vars=c("steps"))
```


## What is mean total number of steps taken per day?
```{r}
dt_total<-dcast(dt_melt, date~variable, sum)
hist(dt_total$steps, main="Histogram of total steps for each day", xlab="Total Steps")

dt_mean <-dcast(dt_melt, date~variable, mean)
dt_median <-dcast(dt_melt, date~variable, median, fill=0)
data.frame(date=dt_mean$date, mean=dt_mean$steps, median=dt_median$steps)
```

## What is the average daily activity pattern?
```{r}
dt_mean <-dcast(dt_melt, interval~variable, mean, na.rm=TRUE)
plot(dt_mean$interval, dt_mean$steps, type='l', xlab="Interval", ylab="Mean Steps")

# Find interval with maximum number of steps
dt_mean[dt_mean$steps==max(dt_mean$steps),]
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  
**Interval 835**.

## Imputing missing values
The missing values are being replaced with the mean number of steps at that 5-min interval for all days.

```{r}
# No. of rows with NA
length(dt[is.na(dt$steps),][,1])

#dt_mean <-dcast(dt_melt, date~variable, mean)
dt2 <- read.csv("activity.csv")

# Replacing NAs with the mean number of steps of the interval for all days.
for (x in 1:length(dt2[,1])) {
  interval <- dt2[x,3]
  if (is.na(dt2[x,1])) { 
    interval_mean <- dt_mean[dt_mean$interval==interval,][2]
    dt2[x,1] <- as.integer(interval_mean)
  }
}

dt_melt2 <- melt(dt2, id=c("date", "interval"), meas.vars=c("steps"))
dt_total2<-dcast(dt_melt2, date~variable, sum)
par(mfrow=c(1,2))
hist(dt_total$steps, main="With NA", xlab="Total Steps")
hist(dt_total2$steps, main="Imputing NA", xlab="Total Steps")

dt_mean2 <-dcast(dt_melt2, date~variable, mean)
dt_median2 <-dcast(dt_melt2, date~variable, median, fill=0)
data.frame(date=dt_mean2$date, mean=dt_mean2$steps, median=dt_median2$steps)
```

Imputing NA values with the mean value of 5-min interval does not change the overall shape of the histograms 
before and after imputing.

## Are there differences in activity patterns between weekdays and weekends?
```{r cache=TRUE}
dt2$day<-lapply(dt2$date, FUN=function(x){
    if(length(grep("Sat|Sun", weekdays(as.Date(x))))){
      "Weekends"
    } else {
      "Weekdays"
    }
})
dt2$day <- factor(dt2$day, levels=c("Weekdays", "Weekends"))
dt2_melt <- melt(dt2, id=c("day", "interval", "date"), meas.vars=c("steps"))
dt2_mean<-dcast(dt2_melt, interval+day~variable, mean)

library(lattice)
xyplot(steps~interval|day, data=dt2_mean, type='l', layout=c(1,2), ylab="Number of Steps")
```

During weekends, activities are more spread out throughout the days. 
During weekdays, there are more activities in the early morning and peak at around 0835h.

